### 2018-06-17 CS761-RandomizedAlgorithms/assignments/a2.tex
q4

@@

$$Pr(X_{i, <} =1) = (\frac{n}{2} -k -1) / n = \frac{1}{2} - \frac{k+1}{n}$$

@@

$$E[X_<] = \sum_{i\in L'} Pr(X_{i,<} = 1)  = \sum_{i \in L'} (\frac{1}{2} - \frac{k+1}{n}) = \frac{m(n-2(k+1))}{2n}.$$

@@

$$Pr(X_< \geq \frac{m}{n}) = Pr(X_< \geq (1+\delta) E[X_<])$$
where $$\delta = \frac{2}{\frac{n-2(k+1)}{2}}-1=\frac{4-(n-2(k+1))}{n-2(k+1)}$$

@@

%Problem 5
\newpage
\section{}
\end{document}

>>>>

I discussed this problem with Sharat.

@@

$$Pr(X_{i, <} =1) = (\frac{n}{2} -k -1) / n = \frac{n-2k-2}{2n}$$

@@

$$E[X_<] = \sum_{i\in L'} Pr(X_{i,<} = 1)  = \sum_{i \in L'} \frac{n-2k-2}{2n} = \frac{m(n-2(k+1))}{2n}.$$

@@

$$Pr(X_< \geq \frac{m}{2}) = Pr(X_< \geq (1+\delta) E[X_<])$$
where $$\delta =\frac{2(k+1)}{n-2(k+1)}.$$
We proceed by assuming $k < \frac{n}{4}-1$ so that $0<\delta <1$. Ultimately we are interested in results on how many samples $m$ we need to take to obtain a certain quality of approximation, and so bounding $k$ is reasonable.
Since $0 < \delta < 1$, and $X_<$ is a sum of independent random coin flips, we can apply the second Chernoff bound from Theorem on page $4$ of Lecture $2$, obtaining
$$Pr(X_< \geq \frac{m}{2}) < \exp(\frac{-\delta^2E[X_<]}{3}) = \exp(\frac{-2(k+1)^2m}{3n(n-2(k+1))}).$$
Hence the probability of failure is at most
$$2\exp(\frac{-2(k+1)^2m}{3n(n-2(k+1))})$$
To succeed with probability at least $0.9999$ we want this value to be at most $10^{-5}$. So it will suffice to choose $m$ and $k$ to satisfy
$$\exp(\frac{-2(k+1)^2m}{3n(n-2(k+1))}) \leq 10^{-6}$$
Rearranging yields the expression
$$(k+1)^2m \geq 9\ln(10) n^2 - 18\ln(10) n(k+1)$$
Dropping the non-positive term on the left-hand side, as this can only improve our probability of success, gives us the following expression for the trade-off between $m$ and $k$:
$$(k+1)^2m = 9\ln(10)n^2 = \Omega(n^2).$$
If we want $k \leq \epsilon n$ for small $\epsilon$ (say $<\frac{1}{4}$) then we need $m$ to satisfy
$$(\epsilon n+ 1)^2 m = 9\ln(10)n^2$$
i.e. 
$$m = \frac{9\ln(10)n^2}{(\epsilon n + 1)^2}.$$
Similarly if we want $k \leq n^{1-\eta}$ for $\eta \leq \frac{1}{2}$ then we take (again assuming desired $k$ is less than $\frac{1}{4}n-1$)
$$m =  \frac{9\ln(10)n^2}{(n^{1-\eta} + 1)^2}.$$

@@

We assume that the determinant of an $n \times n$ matrix where each element is an entry of $\F[x]$ of degree at most $d$ can be computed in $O(dn^\omega)$ field operations.
\subsection{a}
Let $G=(U\cup V, E)$ be a bipartite graph where each edge is coloured either red or blue. We may assume $|V| = |U|$ by adding dummy vertices as needed. Let $k \in \Z$ be such that $1 \leq k \leq |E|$. We construct an $n\times n$ matrix $A$, where $|A| = n$, of indeterminants as follows: for every pair of vertices $u \in U$, $v\in V$ set
$$A_{uv} = \begin{cases}
x_{uv}, \text{if $uv$ is in $E$ and is coloured blue} \\
rx_{uv}, \text{if $uv$ is in $E$ and is coloured red} \\
0, \text{otherwise}.
with indeterminants $x_{uv}$ for each edge $uv$ and indeterminant $r$. For a perfect matching $M$ on $G$ define $s(M)$ to be the sign of the permutation that sends $i$ to $j$ if $ij \in M$. Also define $c_M$ to be the number of edges coloured red in $M$. Let $\cM$ be the set of perfect matchings on $G$. Then we observe that
$$\det(A) = \sum_{\text{permutations } \sigma}\text{sign}(\sigma) \Pi_{i=1}^n A_{i,\sigma(i)} = \sum_{M \in \cM} s(M) r^{c_M}\Pi_{uv \in M}x_{uv}.$$
Let $\cM_\ell$ be the set of perfect matchings on $G$ with $\ell$ red edges. Then we can rearrange the determinant of $A$ as
$$\det(A) = \sum_{\ell = 1}^n (\sum_{M \in \cM_\ell} s(M)\Pi_{uv \in M}x_{uv})r^\ell.$$
As in the proof of Edmonds' theorem from Lecture $8$ the coefficient of the term corresponding to $r^k$ is not identically zero if and only there is a perfect matching in $G$ with $k$ red edges. This coefficient is a polynomial in variables $x_{uv}$ over $uv \in E$, of degree $n$. So if we substitute values into the variables $x_{uv}$, $uv \in E$, drawn independently and uniformly at random from a finite field $\F$ then by the Schwartz-Zippel Lemma this coefficient will be $0$ with probability at most $\frac{n}{|\F|}$. Choose $\F$ so that this probability is as low as desired.
Then after choosing random values for entries $x_{uv}$, $A$ becomes a matrix whose entries are in $\F[r]$ with degree at most $1$. So in $O(n^\omega)$ field operations we can compute $\det(A)$. Then we query the coefficient of the $k$-th term, which will always be $0$ if there are no perfect matchings with $k$ red edges, and with high probability will be non-zero if there is a perfect matching with $k$ red edges. $\blacksquare$.
\subsection{b}
Now we suppose that each edge $uv \in E$ has an associated weight $w_{uv}$, and the maximum weight of an edge is $W$. We use a similar idea to the previous problem, this time creating our $n\times n$ matrix $A$ as follows: for each pair $u \in U$, $v\in V$ set
$$A_{uv} = \begin{cases}
	x_{uv}r^{w_{uv}}, &\text{if $uv \in E$} \\
	0, &\text{otherwise.}
Let $w^*$ be the maximum weight of a perfect matching on $G$. Let $\cM^\ell$ be the set of perfect matchings on $G$ of weight $\ell$. Now $\det(A)$ is a polynomial of the form
$$\det(A) = \sum_{\ell = 1}^{w^*} (\sum_{M \in \cM^\ell} s(M)\Pi_{uv \in M}x_{uv})r^\ell.$$
As previously argued, if we choose large enough $\F$ then drawing the values for $x_{uv}$, $uv \in E$, from $\F$ uniformly independently at random, then by Schwartz-Zippel the coefficient on $r^{w^*}$ in the determinant polynomial will be non-zero with high probability. After assigning values to each $x_{uv}$, $A$ is matrix whose entries are in $\F[r]$ with degree at most $W$. So we can compute $\det(A)$ in O$(Wn^\omega)$ field operations. Querying the degree of the leading term in the resulting polynomial will give the maximum weight of a perfect matching on $G$ with high probability. $\blacksquare$
%Problem 5
\end{document}
